---
title: "DATA 607 Project 1"
author: "Zach Dravis"
date: "2/24/2018"
output: html_document
---
# DATA 607 Project 1
For this project, I was given a text file with chess tournament results where the information has some structure (See https://github.com/zachdravis/CUNY-DATA-607/blob/master/Project%201/DATA%20607%20Project%201%20Data.txt).

The task was to create an R Markdown file that generates a .CSV file with the following information for all of the players:
- Player’s Name
- Player’s State
- Total Number of Points
- Player’s Pre-Rating
- Average Pre Chess Rating of Opponents

For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and
dividing by the total number of games played.

What follows is my documented methodology so that others may learn from this example of manipulating data in text files.

## Import Data and Load Libraries
To download data please visit: https://github.com/zachdravis/CUNY-DATA-607/blob/master/Project%201/DATA%20607%20Project%201%20Data.txt
```{r}
library(stringr)

data <- read.csv("/Users/zachdravis/Documents/GitHub/CUNY-DATA-607/Project 1/DATA 607 Project 1 Data.txt")
```

## Clean Data
Here I eliminate the dash delineators and also make it so that each record / player only has one row of data--in line with concepts of tidy data.

I'll use the sequence function to do so, since these data are uniformly structured
```{r}
data <- data.frame(data[-c(seq(3, 196, by =3)), ], stringsAsFactors = F)

NewData <- data.frame(Col1 = paste(data[seq(3, 129, by = 2),], data[seq(4,130, by = 2),], sep=""), stringsAsFactors = F)
```

So now each individual chess player has their own row.

## Given Example
Here I just create a data frame with the given example for reference.
```{r}
Example <- data.frame(PlayerName = "Gary Hua", PlayerState = "ON", TotalPoints = 6.0, PreRating = 1794, AvgPreChessRatingOfOpponents = 1605)
```

## Add Player Number and split the strings
```{r}
NewData$PlayerNumber <- 1:64

#Create a function to split the information
SplitPlayerInfo <- function(x){
  PlayerSplit <- strsplit(NewData[x,1], "\\|")
  return(PlayerSplit)
}

#Make a master list of all the other lists
MasterList <- NULL
for(i in 1:64){
  name <- paste("Player", i, sep = "")
  x <- assign(name, SplitPlayerInfo(i))
  MasterList <- append(MasterList, x)
}

```

##Player Name
Here I extract player name from the string and then append it to a vector that I add to the data frame.
```{r}
AllPlayerNames <- NULL
for(i in 1:64){
  PlayerName <- MasterList[[i]][2]
  x <- str_c(unlist(str_extract_all(PlayerName, "[:alpha:]+\\b")), collapse = " ")
  AllPlayerNames <- append(AllPlayerNames, x)
}

NewData$PlayerName <- AllPlayerNames
```

##Player's State
Same technique as above.
```{r}
AllPlayerStates <- NULL
for(i in 1:64){
  PlayerState <- MasterList[[i]][11]
  x <- str_extract(PlayerState, "[:alpha:]{2}")
  AllPlayerStates <- append(AllPlayerStates, x)
}

NewData$PlayerStates <- AllPlayerStates
```

##Player's Total Points
Same technique used for name and state.
```{r}
AllPlayerTotalPoints <- NULL
for(i in 1:64){
  PlayerPoints <- MasterList[[i]][3]
  x <- str_extract(PlayerPoints, "\\d\\.\\d*")
  AllPlayerTotalPoints <- append(AllPlayerTotalPoints, x)
}

NewData$TotalPoints <- AllPlayerTotalPoints
```

##PreRating
Same technique for the above variables including removigng the "P.."
```{r}
AllPlayerPreRating <- NULL
for(i in 1:64){
  PlayerPreRating <- MasterList[[i]][12]
  x <- unlist(str_extract_all(PlayerPreRating, "[[:digit:]P]+\\b"))
  x <- x[2]
  x <- str_replace(x, "P.*", "")
  AllPlayerPreRating <- append(AllPlayerPreRating, x)
}

NewData$PreRating <- AllPlayerPreRating
```

##Pre Tournament Rating of Opponents
Here I use two nested for loops to calculate the average pre tournament rating for each individual's opponents.
```{r}
OpponentRatings <- NULL
AllPlayersAvgOpponentRatings <- NULL
for(i in 1:64){
  Opponents <- MasterList[[i]][4:10]
  Opponents <- str_extract(Opponents, "\\d*$")
  Opponents <- Opponents[Opponents != ""]
  Opponents <- as.numeric(Opponents)
  for(j in Opponents){
    Matches <- NewData$PlayerNumber == j
    OpponentRatings <- append(OpponentRatings, as.numeric(NewData$PreRating[Matches]))
  }
    AllPlayersAvgOpponentRatings <- append(AllPlayersAvgOpponentRatings, (sum(OpponentRatings) / length(OpponentRatings)))
    OpponentRatings <- NULL
}

NewData$AvgOpponentRating <- AllPlayersAvgOpponentRatings
```

#Drop the now unneded column and rename dataframe
```{r}
ChessPlayerData <- NewData
ChessPlayerData <- ChessPlayerData[, -c(1)]
```

For more information on manipulating strings, see the soon to be posted stringr guide here: https://github.com/zachdravis/R_Guides/tree/master